Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}

### 2017-09-23 paper/abstract.tex
added abstract

Decisions on how to best operate large complex plants such as natural gas processing, oil refineries, and energy efficient buildings are becoming ever so complex that model-based predictive control (MPC) algorithms must play an important role. However, a key factor prohibiting the widespread adoption of MPC, is the cost, time, and effort associated with learning first-principles dynamical models of the underlying physical system. An alternative approach is to employ learning algorithms to build black-box models which rely only on real-time data from the sensors. Machine learning is widely used for regression and classification, but thus far data-driven models have not been used for closed-loop control. We present novel Data Predictive Control (DPC) algorithms that use Regression Trees and Random Forests for receding horizon control. 
%DPC also utilizes Gaussian Processes for smart experiment design by recommending how the accuracy of the current black-box models can be improved. 
We demonstrate the strength of our approach with a case study on a bilinear building model identified using real weather data and sensor measurements. 
In a one-to-one comparison, we show that DPC explains 70\% variation in the MPC controller.
%We validate our results by a one-to-one comparison against a benchmark MPC controller. 
We further apply DPC to a large scale multi-story EnergyPlus building model to curtail total power consumption in a Demand Response setting. 
In such cases, when the model-based controllers fail due to modeling cost, complexity and scalability, our results show that DPC curtails the desired power usage with high confidence.

>>>>

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Data-driven models have been intensively studied in the last year to overcome this problem. However, closed-loop control systems based on such models have not received much attention. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC, and apply it to the Demand Response problem of a building and to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
